Manifold Clustering in the Setting of Generalized Random Dot Product Graphs

### 2023-05-24 sdss-2023/poster-2.Rmd
minor fix to poster
minor changes
minor updates

1. Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.
2. Initialize community labels $z_1, ..., z_n$.
3. While change in the loss function $L = \sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$: 
    i. For $k = 1, ..., K$:
        a. Define $X_k$ as the rows of $X$ for which $z_i = k$.
        b. Fit curve $g_k$ and positions $t_{k_i}$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.
    ii. For $k = 1, ..., K$: 
        a. Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.

@@

**Theorem**. 
Let an MBM be such that the manifolds are described as polynomial curves of order $R$. 
Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
Then $K$-curves clustering outputs estimates such that  
$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; X) \stackrel{p}{\to} 0.$$

@@

@@

qgraph::qgraph(A, 
               groups = factor(z), 
               legend = FALSE,
               mar = c(0, 0, 0, 0))

@@

ggplot(clust.summary.df) + 
  theme_bw() + 
  theme(text = element_text(size = 10)) + 
  scale_y_log10() +
  scale_x_log10(breaks = n.vec) +
  labs(y = 'community detection error rate',
       x = 'n',
       colour = 'known labels per community', 
       shape = NULL) + 
  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_errorbar(aes(x = n, 
                    ymin = first.q.rate, ymax = third.q.rate, 
                    colour = factor(nsamp)),
                width = .125) + 
  theme(legend.position = 'bottom') + 
  scale_colour_brewer(palette = 'Set1')
```

>>>>

1. Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.
2. Initialize community labels $z_1, ..., z_n$.
3. While change in the loss function $L = \sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$: 
    i. For $k = 1, ..., K$:
        a. Define $X_k$ as the rows of $X$ for which $z_i = k$.
        b. Fit curve $g_k$ and positions $t_{k_i}$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.
    ii. For $k = 1, ..., K$: 
        a. Assign $z_i \leftarrow \arg\min_k \|x_i - g_k(t_i)\|^2$.

@@

**Theorem**. 
Let an MBM be such that the manifolds are described as polynomial curves of order $R$. 
Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
Then $K$-curves clustering outputs estimators such that  
$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; A) \stackrel{p}{\to} 0.$$

@@

degs <- colSums(A)
A.trunc <- A[degs > 0, degs > 0]
z.trunc <- z[degs > 0]
# X <- X[degs > 0, ]
# Xhat <- Xhat[degs > 0, ]

@@

qgraph::qgraph(A.trunc, 
               groups = factor(z.trunc), 
               legend = FALSE,
               mar = c(0, 0, 0, 0))

@@

ggplot(clust.summary.df) + 
  theme_bw() + 
  theme(text = element_text(size = 10)) + 
  scale_y_log10() +
  scale_x_log10(breaks = n.vec) +
  labs(y = 'error rate',
       x = 'n',
       colour = 'known labels per community', 
       shape = NULL) + 
  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_errorbar(aes(x = n, 
                    ymin = first.q.rate, ymax = third.q.rate, 
                    colour = factor(nsamp)),
                width = .0625) + 
  theme(legend.position = 'bottom') + 
  scale_colour_brewer(palette = 'Set1')
```

### 2023-05-24 sdss-2023/slides.Rmd
minor updates

Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$. 
Define probability distribution $F$ with support $[0, 1]^r$. 
Then the following mixture model is a *manifold block model*:

>>>>

Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \subset \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$. 
Define probability distribution $F$ with support $[0, 1]^r$. 
Then the following mixture model is a *manifold block model*:
